---
title: "TE Chapter 16 Doc"
output: html_document
date: "2023-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How Does It Work?\
> 1.	 You observe the number of vacations taken by Zac and Skylar in 2012, 2013, and 2014. In those years, Zac took 3, 7, and 5 vacations, respectively. Skylar took 2, 6, and 10. 

**a.	Isolate the numbers that represent the variation between Zac and Skylar in their vacation-taking.**\
```{r}
((3+7+5)/3) 
((2+6+10)/3)
```
  To look at the variation between Zac and Skylar you would compare 5 (Zac) and 6 (Sylar).
**b.	Isolate the variation within Zac and within Skylar in their vacation-taking.**\
```{r}
# Zac 
3 - 5
7 - 5
5 - 5
# Skylar
2 - 6
6 - 6
10 - 6
```
Zac's within variation in vacation-taking compares the fact that in 2012 he took 2 less vacations than average, in 2013 he took 2 more vacations than average, and in 2014 he took an average amount of vacations. Skylar's within variation in vacation-taking compares the fact that in 2012 she took 4 less vacations than average, in 2013 she took an average number of facations, and in 2014 she took four more vacations than average. \

**c.	(Difficult!) We perform a fixed effects analysis of the effect of vacations on happiness. A vacation increases Zac’s happiness by 1 “happiness point,” but it increases Skylar’s happiness by 2 “happiness points.” Will our fixed effects estimate likely give us an answer closer to 1, closer to 2, or exactly 1.5?**\

Not really sure how to work through the reasoning for this one. I am thinking that beacause her within variation is double that the fixed effect will have some larger value than 1. I am going to guess it is closer than 2, but I am not quite sure how to think through this. \

> 2.	You are interested in the effect of cultural events on the levels of trust in a city. Perhaps big events like concerts bring people together and they can trust each other more. You plan to look at the relationship between trust and number of events in a given year, with fixed effects for city. Draw a causal diagram for this research question with at least four back door paths. Which paths will be closed by fixed effects, and which will remain open?

My four variables are crime levels (C), the money of the city (M), diversity (D), and history of the ethnic groups in the area (H). I will label trust as T and cultural event as CE. 
```{r}
library(ggdag)
library(tidyverse)

dag_coords <-
  tibble(
    name = c("C", "M", "D", "H", "CE", "T"),
    x = c(2, 5, 3.5, 4.5, 1, 8),
    y = c(4,4,6,6,1,1)
  )

p1 <-
  dagify(T ~ CE,T ~ H, CE ~ H, T ~ C, CE ~ C, T ~ C, CE ~ D, CE ~ M, T ~ M,
    coords = dag_coords
  ) %>%
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "blue4", alpha = 1 / 4, size = 10) +
  geom_dag_text(color = "cornflowerblue") +
  geom_dag_edges(edge_color = "blue4") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(0.1, 0.1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(0.2, 0.2)) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(
    title = "DAG for Question 2")

p1
```

This is a hard one for me to concceptualize, because to me all of these things seem to mostly not be constant. Diversity can change even if slowly, the crime rate likely changes more frequently, the money of the city may change yearly (either by what is in the budget that might actually effect. these things or donations given, or changes in taxes). The only thing that I know would not change is the history. However, I am thinking that the things that change more slowly like the money of the city, the diversity in the city, and the history of the city may be able to be closed with fixed effects. \

> 3.	Classify each of the following forms of variation as “between variation”, “within variation”, or a combination of both.

**a.	(Individual = person) How a child’s height changes as they age.** This is within variation. \
**b.	(Individual = person) In a data set tracking many people over many years, the variation in the number of children a person has in a given year.** I think this is a combination of both, there is variation across each person in the number of children they have, but also individuals may have more children each year and thus in Y1 they may have 2 and by Y4 they may have 3.\
**c.	(Individual = city) Overall, Paris, France has more restaurants than Paris, Texas.** This is between variation because it is comparing two different cities. \
**d.	(Individual = genre) The average pop music album sells more copies than the average jazz album** I think because it says "the average" of each its getting rid of within variation... so this is between variation specifically because its two diferent generes.\
**e.	(Individual = genre) Miles Davis’ Kind of Blue sold very well for a jazz album. ** I think this is within variation because its within one genre (jazz) and its not really comparing all of Miles Davis' work per say. \
**f.	(Individual = genre) Michael Jackson’s Thriller, a pop album, sold many more copies than Kind of Blue, a jazz album.**
I think that since this is similar to an average, within variation of each genre is ignored here and this is between variation looking at two different genres. \

> 4.	Why does the process of taking each observation relative to its individual-level mean have the effect of “controlling for individual”?

This is because it takes all of the variation that is for one individual and aggregates it to one number in affect controlling for the individual because the mean gives us a constant value for comparison.\

# How is it Performed? 
1.	You are interested in the effect of cultural events on the levels of trust in a city. You run a regression of trust levels (on a 0-100 scale) on the number of cultural events with city fixed effects and get a coefficient on cultural events of 3.6. Assume that there are still some back doors open, so do not interpret the result causally. Interpret the 3.6, explaining it in an English sentence. 
2.	You are interested in the effect of cultural events on the levels of trust in a city. You run a regression of trust levels (on a 0-100 scale) on the number of cultural events with city and year fixed effects and get a coefficient on cultural events of 2.4. Assume that there are still some back doors open, so do not interpret the result causally. Interpret the 2.4, explaining it in an English sentence.
3.	Two-way fixed effects with terms for both individual and time are often referred to as “controlling for individual and time effects”. Why might a researcher want to do this rather than just taking individual fixed effects and adding a linear/polynomial/etc. term for time?
4.	Which of the following explains why random effects is likely to do a better job of estimating the individual-level effects than fixed effects, if its assumptions hold?
a.	Because it makes the assumption that the individual effects are unrelated to the other predictors, which breaks that back door and thus reduces bias.
b.	Because random effects allows some amount of between variation into the model, and some of the real individual effect is that between variation.
c.	Because it uses the information from the entire data set to estimate each individual effect, rather than relying on only a few observations per individual.
d.	It won’t. Enforcing Durbin-Wu-Hausman makes both methods produce the same estimates anyway. 

