---
title: "missing data"
output: html_document
date: "2023-02-07"
---

# Chatper 15: Missing Data\

# 15.1: Measurement Error\

Let's see how to incorporate measurement error into the model.

```{r}
library(rethinking) 
data(WaffleDivorce)
d <- WaffleDivorce
# points
plot( d$Divorce ~ d$MedianAgeMarriage, ylim = c(4,15),
xlab = "Median age marriage",ylab = "Divorce rate")
# standarderrors
for (i in 1: nrow(d)){
ci <-d$Divorce[i] + c(-1,1) * d$Divorce.SE[i]
x <- d$MedianAgeMarriage[i]
lines( c(x,x),ci)
}
```

There is a lot of variation in how uncertain the observed divorce rate (reflected in varying lengths of the vertical line segments).\

### 15.1.1: Error on the outcome

How do we simulate measurement error?\
  1. generate the true values of the variables\
  2. simulate the observatino process (where the measurement error would occur)\
  
  
We have reported standard errors... how can we use this information in the statistical model?\

Here is how to define the error distribution for each divorce rate:\

$$
D_{OBS,i} \sim Normal(D_{TRUE,i},D_{SE,i})\\
D_{TRUE,i} \sim Normal(\mu_i, \sigma)\\
\mu_i = \alpha + \beta_AAi + \beta_MMi\\
\alpha \sim Normal(0,0.2)\\
\beta_A \sim Normal(0,0.5)\\
\beta_M \sim Normal(0,0.5)\\
\sigma \sim Exponential(1)\\
$$
This is like a linear regression but with the addition of the top line which connects the observation to the true value

Let's put it into code

```{r}
library(tidyverse)
library(dplyr)

# transform variables as needed
dlist <- d %>% 
  mutate(D_obs = standardize(Divorce),
       D_sd = Divorce.SE/sd(Divorce),
       M = standardize(Marriage),
       A = standardize(MedianAgeMarriage),
       N = nrow(d)) 

# create model with d_obs and d_sd

m15.1 <- ulam(
alist(
D_obs ~ dnorm(D_true, D_sd),
vector[N]: D_true ~ dnorm(mu, sigma),
mu <- a + bA * A + bM * M,
a ~ dnorm(0, 0.2),
bA ~ dnorm(0, 0.5),
bM ~ dnorm(0, 0.5),
sigma ~ dexp(1)) , data = dlist, chains = 4, cores = 4)

precis(m15.1, 2)

```
Comparing with the results for this model in chapter 5 and now, we see that including measurement error reduces the association between divorce and age at marriage (bA). \

This is not always the case though, sometimes  measurement error can hide effects (not just exaggerate like in this example).\

###15.1.2: Error on the outcome and predictor. \


$$
D_{OBS,i} \sim Normal(D_{TRUE,i},D_{SE,i})\\
D_{TRUE,i} \sim Normal(\mu_i, \sigma)\\
\mu_i = \alpha + \beta_AAi + \beta_MM_{TRUE,i}\\
M_{OBS,i} \sim Normal(M_{TRUE,i},M_{SE,i})\\
M_{TRUE,i} \sim Normal(0,1)\\
\alpha \sim Normal(0,0.2)\\
\beta_A \sim Normal(0,0.5)\\
\beta_M \sim Normal(0,0.5)\\
\sigma \sim Exponential(1)\\
$$


Let's see it in code.\

```{r}
# using same data from above, add M_obs and M_sd

dlist <- dlist %>% 
  mutate(M_obs = standardize(Marriage),
         M_sd = Marriage.SE/sd(Marriage),)

m15.2 <-ulam(
alist(
D_obs ~ dnorm(D_true, D_sd),
vector[N]: D_true ~ dnorm(mu,sigma),
mu <- a + bA * A + bM * M_true[i],
M_obs ~ dnorm(M_true, M_sd),
vector[N]: M_true ~ dnorm(0,1),
a ~ dnorm(0,0.2),
bA ~ dnorm(0,0.5),
bM ~ dnorm(0,0.5),
sigma ~ dexp(1)
) , data = dlist, chains = 4, cores = 4)
```

Notice that adding error on the predictor didn't change the major inference, but it did provide updated estimates of marriage rates itself. \


# 15.2: Missing Data\

What can be done when there is no measurement at all?\

  1. complete case analysis = drop all cases with any missing values\
        - can produce bias, depnding upon the causal details
  2. replace missing values with some assumed value (like the mean or a reference value like zero)\
    - the model will think we know these values certainly instead of as a guess
  3. think causally about missingness and use the model to impute missing values
    
    
# CLASS NOTES
    
Standard error = uncertainty about the error assuming everything is measured perfectly\
Measurement error = uncertainty in the measurement\
  
  
